Analysis of observational data, to calculate the Delayed-Tau model parameters

### 2025-03-03 thesis.qmd
start calc chapter

execute: 
  echo: false
  eval: true
  warning: false
editor: visual

@@

![Plot showing the linear relation $\log_{10}\text{SFR}_{FUV} = \log_{10}\text{SFR}_{H\alpha}$, as well as their distributions, based on the data from UNGC](figure/log_SFR_FUV_Ha.png){#fig-SFR-Ha-FUV fig-align="center" width="500"}

@@

$$
\overline{\text{SFR}_\text{del}} = \frac{\zeta M_*}{t_sf}
$$ {#eq-averaged-SFR-M*}

@@

>>>>

execute: 
  echo: false
  eval: true
  warning: false
#editor: visual

@@

![Plot showing the linear relation $\log\text{SFR}_{FUV} = \log\text{SFR}_{H\alpha}$, as well as their distributions, based on the data from UNGC](figure/log_SFR_FUV_Ha.png){#fig-SFR-Ha-FUV fig-align="center" width="500"}

@@

$$
\overline{\text{SFR}_\text{del}} = \frac{\zeta M_*}{t_sf}
$$ {#eq-averaged-SFR-M}

@@

Using the merged catalog, we can use the SFR and $M_*$ data to calculate the parameters for the delayed-$tau$ model, for each individual galaxy. The parameters are going to be calculated using two different computational methods.

Firstly, using the Newton-Ramphson algorithm, the parameters $\tau$ and $A_{del}$ can be calculated from the equations @eq-delayed-current and @eq-delayed-current-averaged, assuming the same $t_{sf} = 13.6$ Gyr for every galaxy, meaning that all the galaxies start the star formation proccess around the same time, 13.6 Gyr ago, or 0.2 Gyr after the Big Bang. This is the same method used in @haslbauerCosmologicalStarFormation2023, but instead of using only the SFR from UNGC and adopting a mass-to-light ratio of 0.6 from the $L_K$, we use the merged catalog.

The second method for the calclulation of the parameters is the MCMC method. Using the logarithmic form of the model

$$
\log(\text{SFR})= \log(A_{del})+\log(x)-\log(\tau)-x\cdot\log e,\text{ where } x=\frac{t_sf}{\tau},
$${#eq-log-SFR}

and by combining @eq-delayed-current-averaged and @eq-averaged-SFR-M:

$$
A_{del}=\frac{\zeta M_*}{1-(1+x)e^{-x}},
$${#eq-A}

we can create a MCMC model to calculate the parameters, with the priors:

- $\tau \sim \text{normal(mean=4 Gyr, }\sigma = 1 )$, within the limits $[1,20]$ Gyr,
- $t_{sf}\sim \text{uniform(lower bound = 1 Gyr, upper bound = 13.8 Gyr)}$,
- $\zeta \sim \text{normal(mean=1.3, }\sigma = 0.01)$, within the limits $[1,2]$,

and the likelihood:

- $\log(\text{SFR}_{obs}) \sim \text{normal}\left(\log(\text{SFR}_\text{MCMC}), \sigma = 0.1\right)$
